Kids SUMs Summary

### 2020-01-28 moisture_summary.Rmd
summary stat updates

---
title: "Kids Moisture Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 27 Jan 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
```

@@

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home) %>% 
  ungroup() %>% 
  arrange(area, home)

moisture_filtered <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

@@

>>>>

---
title: "Kids Moisture Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 28 Jan 2020"
output: pdf_document
---

@@

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(readr)
library(knitr)
```

@@

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home, first_datetime_new) %>% 
  separate(first_datetime_new, c("first_date", "first_time"), sep = " ", remove = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home)

moisture_filtered <- moisture_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>%
  arrange(area, home)

moisture_filtered_time <- moisture_filtered %>% 
  # fix wrong date
  mutate(moisture_date = as.character(moisture_date),
         moisture_date = if_else(home == "CH220" & moisture_date == "2014-04-10",
                                 "2017-04-10", moisture_date),
         moisture_date = ymd(moisture_date)) %>% 
  mutate(moisture_date = ymd(moisture_date),
         first_date = ymd(first_date),
         time_diff = moisture_date - first_date,
         time_diff = as.numeric(abs(time_diff))) %>% 
  group_by(home) %>% 
  mutate(time_diff_rank = if_else(time_diff == min(time_diff, na.rm = TRUE), 1, 0)) %>%
  ungroup() %>% 
  filter(time_diff_rank == 1)
```

@@

# Summarize numeric vars
kids_moisture <- moisture_filtered_time %>%  
  group_by(area) %>% 
  summarize("Mean" = mean(time_diff, na.rm = TRUE), 
            "SD" = sd(time_diff, na.rm = TRUE),
            "min" = min(time_diff, na.rm = TRUE), 
            "med" = median(time_diff, na.rm = TRUE), 
            "max" = max(time_diff, na.rm = TRUE))
kable(kids_moisture, digits = 1, align = "c")

### 2020-01-28 pm_summary.Rmd
summary stat updates

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 23 Jan 2020"
output: pdf_document
---

@@

# Summary stats
```{r}
pm_summary <- pm_clean %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(pm < 500) %>% 
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "Median PM" = median(pm, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
  select(-area) %>% 
  t()

kable(pm_summary, col.names = c("Alaska", "Navajo", "Montana"), 
      digits = 1, align = "c")
```

>>>>

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 28 Jan 2020"
output: pdf_document
---

@@

# Summary stats
```{r}
pm_summary <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(pm < 500) %>% 
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
  select(-area) %>% 
  t()

kable(pm_summary, col.names = c("Alaska", "Navajo", "Montana"), 
      digits = 1, align = "c")

pm_by_day <- pm_clean %>% 
  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
  mutate(day_of_week = factor(day_of_week,
                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                            "Friday", "Saturday", "Sunday"),
                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(pm < 500) %>% 
  group_by(area, sampling_month) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75),
            "Max PM" = max(pm, na.rm = TRUE))
pm_by_day
```

### 2020-01-28 sums_summary.Rmd
summary stat updates

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 23 Jan 2020"
output: pdf_document
---

@@

>>>>

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 28 Jan 2020"
output: pdf_document
---

@@

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home, pm_datetime_new) %>% 
  separate(pm_datetime_new, c("sample_date", "sample_time"), sep = " ", remove = TRUE) %>% 
  mutate(sample_date = ymd(sample_date)) %>% 
  ungroup() %>% 
  arrange(area, home)

sums_filtered <- sums_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)

sums_filtered_date <- sums_clean %>% 
  mutate(sample_date = date_sums,
         sample_date = ymd(sample_date)) %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id", "sample_date")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Function for calculating heating events using specified threshold temps
## Also specify the "gap" time between events
```{r}
thresholds_function <- function(data, temp, gap) {

thresholds <- data %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event" | threshold_temp_event == "start") %>% 
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event = lead(event),
         time_diff_min = (event - start),
         time_diff_min = as.numeric(time_diff_min/60)) %>% 
  filter(!is.na(time_diff_min)) %>% 
  mutate(next_event = (lead(start)-event),
         next_event = as.numeric(next_event)/3600,
         event2 = if_else(next_event <= gap, lead(event), event),
         time_diff_min2 = (event - start),
         time_diff_min2 = as.numeric(time_diff_min2/60),
         event3 = if_else(next_event <= gap & lead(next_event) <= gap, lead(event, 2), event2),
         time_diff_min3 = (event3 - start),
         time_diff_min3 = as.numeric(time_diff_min3/60)) %>% 
  distinct(event3, .keep_all = TRUE) %>% 
  mutate(time_diff_min4 = (event3 - start),
         time_diff_min4 = as.numeric(time_diff_min4/60)) %>% 
  filter(!is.na(time_diff_min4)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(time_diff_min4)) %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "SD events per day" = round(sd(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0))
thresholds

}

thresholds_function(sums_filtered_date, 25, 1.5)
thresholds_function(sums_filtered_date, 30, 1.5)
thresholds_function(sums_filtered_date, 35, 1.5)
thresholds_function(sums_filtered_date, 40, 1.5)
thresholds_function(sums_filtered_date, 45, 1.5)
thresholds_function(sums_filtered_date, 50, 1.5)
thresholds_function(sums_filtered_date, 55, 1.5)
```

# Function to look at percentage of time spent above a threshold
```{r}
thresholds_percent <- function(data, temp) {

thresholds <- data %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

}

thresholds_percent(sums_filtered_date, 25)
thresholds_percent(sums_filtered_date, 30)
thresholds_percent(sums_filtered_date, 35)
thresholds_percent(sums_filtered_date, 40)
thresholds_percent(sums_filtered_date, 45)
thresholds_percent(sums_filtered_date, 50)
thresholds_percent(sums_filtered_date, 55)
```
